---
title: "Accuracy"
author: "Alene Onion"
date: "October 3, 2018"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---

```{r setuA, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#file.copy("sections/images", ".", recursive = TRUE)
```

##Accuracy Assessment
Accuracy fail: designated 'spike' samples are spiked with a known concentration. The difference between the concentration before and after should be the spiked amount +/- the accuracy criteria set forth in table 3 (screening) and table 4 (routine) 

Also check that the lab has met their own lab accuracy criteria which is in the case narrative of the data package.




%%%!!!%%% I had to replace all the blank cells with NA in 'qc_spike_status' and 'qc_dup_spike_status' to make this work  %%%!!!%%%


%%%!!!%%% This script is inapropriately examining the lab's MS not our MS %%%!!!%%%
- The way I figured this out is that i found that there were MSDs in our data whereas we shouldn't have any MSDs done to our spike samples
- Torganics are measured so low that rather than do a duplicate, they spike the duplicate and produce a spike and a duplicate spike
- our spikes do asses





### Failures per project
```{r message=FALSE, warning=FALSE, results='asis'}
#create pivot table from this table
library(reshape2)
#failure rate per parameter
table<-dcast(spike,project~pass)
table$percFail<-(signif(((table$fail)/(table$fail+table$pass)),2))*100
#table$percFail<-(signif(((table$ms_OR_msd_fail+table$ms_AND_msd_fail)/(table$ms_OR_msd_fail+table$ms_AND_msd_fail+table$pass)),2))*100
table<-table[order(-table$percFail),]
#print table
knitr::kable(table)
```

### Failures per parameter
```{r message=FALSE, warning=FALSE, results='asis'}
#read routine data
#create pivot table from this table
library(reshape2)
#failure rate per parameter
table<-dcast(spike,chemical_name~pass)
table$percFail<-(signif(((table$fail)/(table$fail+table$pass)),2))*100
#table$percFail<-(signif(((table$ms_OR_msd_fail+table$ms_AND_msd_fail)/(table$ms_OR_msd_fail+table$ms_AND_msd_fail+table$pass)),2))*100
table<-table[order(-table$percFail),]
#print table
knitr::kable(table)
```

### Failures: parameter by project
```{r message=FALSE, warning=FALSE, results='asis'}
#read routine data
#create pivot table from this table
spike$chemproject<-paste(spike$chemical_name,spike$project,sep="_")
library(reshape2)
library(knitr)
#failures broken down by project
table<-dcast(spike,chemproject~pass)
#table<-table[table$fail != 0,]
table$percFail<-(signif(((table$fail)/(table$fail+table$pass)),2))*100
#table$percFail<-(signif(((table$ms_OR_msd_fail+table$ms_AND_msd_fail)/(table$ms_OR_msd_fail+table$ms_AND_msd_fail+table$pass)),2))*100
#create separate parameter and project columns
table$parameter<-table$chemproject
table$parameter<-gsub("_.*","",table$parameter)
table$project<-table$chemproject
table$project<-gsub(".*_","",table$project)
table<-unique(table[c('parameter','project','pass','fail','percFail')])
library(DT)
datatable(table, filter = 'top')
```

### Sortable Table
```{r message=FALSE, warning=FALSE, results='asis'}
#read routine data
#create sortable table
#first truncate the table
spike<-unique(spike[c('project','chemical_name','sample_date','sample_name','pass','qc_spike_recovery')])
library(DT)
datatable(spike, filter = 'top')